---
title: "roc_test(fisher)"
author: "Bhummanat"
date: "2024-11-06"
output: html_document
---

This script is used to evaluate the association between GT and other covariates.
This is for troubleshooting purposes.
All tests are performed only for SNPs that passed FDR test.
All tests are performed on a per SNPs position basis.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(broom)
library(ggpubr)
```

```{r Directories}
input_dir <- '/standard/cphg-RLscratch/syv3ap/rotation/TCGA_LGG_imputed_germline/'
```

Some manipulation so that the test could run
Somatic_mut was changed to log to better visualize
```{r }
LGG_allele_dat_sig <- readRDS(paste0(input_dir, 'LGG_allele_dat_sig.rds')) |> 
  mutate(GT = as.integer(as.character(GT)),
         somatic_mut = log(somatic_mut + 1)
         )
LGG_allele_dat_sig$histological_grade <- factor(LGG_allele_dat_sig$histological_grade, ordered = FALSE)
covariate_sig <- readLines(paste0(input_dir, 'covariate_generalized.txt')) |> 
  str_trim() |> 
  setdiff('GT')
```

Perform chisq-test between GT and significant covariates that are categorical
```{r chisq}
chisq_per_snp_results <- LGG_allele_dat_sig |>
    group_by(position) |>
    reframe(
        covariate = covariate_sig,
        p_value = map_dbl(covariate_sig, function(cov) {
            data <- cur_data()  # Get the current group's data
            if (is.factor(data[[cov]]) || is.character(data[[cov]])) {
                table <- table(data[[cov]], data$GT)
                result <- chisq.test(table, simulate.p.value = TRUE, B = 10000)
                result$p.value
            } else {
                NA  # Assign NA for non-categorical covariates
            }
        })
    )
```

This box plot is replaced with violin plot in the chunk below.
Kept as this might still be relevant.
```{r box plot}
# List to store plots
plot_list <- list()

# Loop over each SNP and continuous covariate
unique_snps <- unique(LGG_allele_dat_sig$position)
for (snp in unique_snps) {
  # Filter data for the specific SNP
  snp_data <- LGG_allele_dat_sig |> filter(position == snp)
  
  for (covariate in covariate_sig) {
    # Check if the covariate is continuous
    if (is.numeric(snp_data[[covariate]])) {
      # Remove non-finite values for the current covariate
      plot_data <- snp_data |> filter(is.finite(.data[[covariate]]))
      
      # Create box plot for each SNP and covariate
      p <- ggplot(plot_data, aes(x = factor(GT), y = .data[[covariate]])) +
        geom_boxplot() +
        labs(
          title = paste(covariate, ' ', snp),
          x = "Genotype (GT)",
          y = covariate
        ) +
        theme_minimal()
      
      # Store the plot
      plot_list[[paste(snp, covariate, sep = "_")]] <- p
    }
  }
}

# Save or display all plots
pdf("box_plots_per_snp_covariate.pdf")
for (plot in plot_list) print(plot)
dev.off()
```

Using violin plot and ggpubr to display statistics, labels, and make better for visualization
```{r violin plot}
# Define the list of comparisons for statistical tests
my_comparisons <- list(c("0", "1"), c("0", "2"), c("1", "2"))

# List to store plots (violin)
plot_list <- list()

# Loop over each SNP and continuous covariate
unique_snps <- unique(LGG_allele_dat_sig$position)
for (snp in unique_snps) {
  # Filter data for the specific SNP
  snp_data <- LGG_allele_dat_sig |> filter(position == snp)
  
  for (covariate in covariate_sig) {
    # Check if the covariate is continuous
    if (is.numeric(snp_data[[covariate]])) {
      # Remove non-finite values for the current covariate
      plot_data <- snp_data |> filter(is.finite(.data[[covariate]]))
      
      # Create violin plot with ggviolin for each SNP and covariate
      p <- ggviolin(plot_data, x = "GT", y = covariate, fill = "GT",
                    palette = c("#00AFBB", "#E7B800", "#FC4E07"),
                    add = "boxplot", add.params = list(fill = "white")) +
        labs(
          title = paste(covariate, ' ', snp),
          x = "Genotype (GT)",
          y = covariate
        ) +
        stat_compare_means(comparisons = my_comparisons, label = "p.signif") +  # Add significance levels
        stat_compare_means() +  # Add overall significance
        theme_minimal()
      
      # Store the plot
      plot_list[[paste(snp, covariate, sep = "_")]] <- p
    }
  }
}

# Save or display all plots
pdf("violin_plots_per_snp_covariate.pdf")
for (plot in plot_list) print(plot)
dev.off()
```

